Comparing Apples and Oranges in Youth Depression Treatments?

### 2024-08-12 multilevel_metaregression/manuscript_with_multilevel_metan.qmd
added sensitivity analysis and sex meta-analysis with all studies

###########Libraries needed
library(tidyverse)
library(truncnorm)
library(metafor)
library(meta)
library(kableExtra)
library(flextable)
library(magrittr)
library(officer)
library(RefManageR)
library(ftExtra)
library(dplyr)
library(broom)
library(effsize)

# For referencing
bib_data <- ReadBib("C:/Users/Raphaelle/Documents/AIM Lab/apples_and_oranges/main_git/apples_oranges/emotion_concepts_paper.bib")

@@

df_appl_v_orange <-readxl::read_excel("C:/Users/Raphaelle/Documents/AIM Lab/apples_and_oranges/main_git/apples_oranges/df_appl_v_orange.xlsx")

@@

df_scale_ranges <- read.csv("C:/Users/Raphaelle/Documents/AIM Lab/apples_and_oranges/main_git/apples_oranges/scale_ranges.csv")
df_appl_v_orange <- full_join(df_appl_v_orange, df_scale_ranges, by = "instrument_name")

@@

df_long_for_metan <- df_long_for_metan %>% 
  mutate(percent_women_std_error = sqrt(product_perc_women/ n_overall ))

@@

df_first_row <- read.csv("C:/Users/Raphaelle/Documents/AIM Lab/apples_and_oranges/main_git/apples_oranges/df_first_row.csv")

@@

```{r, echo=FALSE}
#| label: fig-prisma
#| fig-cap: "PRISMA chart summarising sources of included studies"
#| fig-cap-location: top
knitr::include_graphics("PRISMA.png")
```

@@

```{r, warning=FALSE, message = FALSE, echo=FALSE}
# Start by looking at demographics at baseline
# Currently we have both control and active conditions, we just need each study once

@@

# Now look at baseline severity for overall psy vs med
# Need to calculate the variables first

df_for_baseline_severity <- df_long_for_metan %>% 
  group_by(new_study_id) %>%
  mutate(perc_baseline_overall = sum(perc_baseline_mean * baseline_n) / sum(baseline_n)) %>% 
  mutate(pooled_perc_sd_baseline = sqrt(sum((perc_baseline_sd^2) * (baseline_n - 1)) / sum(baseline_n - 1))) %>% 
  mutate(perc_se_baseline = pooled_perc_sd_baseline / sqrt(n_overall)) %>%  # now also do this for untransformed variable
  mutate(baseline_overall = sum(baseline_mean * baseline_n) / sum(baseline_n)) %>% 
  mutate(pooled_sd_baseline = sqrt(sum((baseline_sd^2) * (baseline_n - 1)) / sum(baseline_n - 1))) %>% 
  mutate(se_baseline = pooled_sd_baseline / sqrt(n_overall))

@@

# clinical sample

@@

@@

df_for_sex_metan <- df_for_sex_metan %>% 
  filter(percent_women != 100) # filter out studies with only women %>%

@@

# Now start looking at age variable

@@

For this analysis, we excluded the two psychotherapy trials which included entirely female samples (Moeini, 2019; Shomaker, 2016). As can be seen in @tbl-baseline_results, psychotherapy trials featured a significantly higher percentage of females when compared to medication trials. On average, samples were `r round(results_met_perc_women_overall$TE_random[results_met_perc_women_overall$subgroup == 1][[2]],2)`% (*SE* = `r round(results_met_perc_women_overall$seTE_random[results_met_perc_women_overall$subgroup == 1][[2]],2)`) female across psychotherapy trials and `r round(results_met_perc_women_overall$TE_random[results_met_perc_women_overall$subgroup == 0][[2]],2)`% (*SE* = `r round(results_met_perc_women_overall$seTE_random[results_met_perc_women_overall$subgroup == 0][[2]],2)`) female across medication trials. Excluding sub-clinical and waitlist control studies yielded similar results.

@@

We conducted a series of sensitivity analyses (all figures in the Supplemental Materials). Excluding waitlist control studies (see @fig-plot-means-no-wl) and sub-clinical studies (see @fig-plot-means-clin) yielded a pattern of results very similar to the overall analyses. Next, we examined the data including only those studies that used the CDRS (see @fig-plot-means-cdrs) or the HAMD (see @fig-plot-means-hamd). Medication control and psychotherapy control conditions remained significantly different, though the small number of studies resulted in less precise estimates of the SMDs. Finally, we showed that different values for the pre-post measure correlation had minimal effect on the estimated outcomes (see @fig-stab-sims).

@@

>>>>

###########Libraries needed
library(tidyverse)
library(truncnorm)
library(metafor)
library(meta)
library(kableExtra)
library(flextable)
library(magrittr)
library(officer)
library(RefManageR)
library(ftExtra)
library(dplyr)
library(broom)
library(effsize)
library(here)

#setting paths using here package
here::here()

# For referencing
bib_data <- ReadBib(here("emotion_concepts_paper.bib"))

@@

df_appl_v_orange <-readxl::read_excel(here("df_appl_v_orange.xlsx"))

@@

df_scale_ranges <- read.csv(here("scale_ranges.csv"))
df_appl_v_orange <- full_join(df_appl_v_orange, df_scale_ranges, by = "instrument_name")

@@

df_long_for_metan <- df_long_for_metan %>% 
  mutate(percent_women_std_error = sqrt(product_perc_women/ n_overall))

@@

df_first_row <- read.csv(here("df_first_row.csv"))

@@

```{r, echo=FALSE}
#| label: fig-prisma
#| fig-cap: "PRISMA chart summarising sources of included studies"
#| fig-cap-location: top
knitr::include_graphics(here("PRISMA.png"))
```

@@

# Start by looking at demographics at baseline
# Currently we have both control and active conditions, we just need each study once

@@

#A. Baseline severity meta-analysis--------------

# Overall psy vs med

# Need to calculate the variables first
df_for_baseline_severity <- df_long_for_metan %>% 
  group_by(new_study_id) %>%
  mutate(perc_baseline_overall = sum(perc_baseline_mean * baseline_n) / sum(baseline_n)) %>% 
  mutate(pooled_perc_sd_baseline = sqrt(sum((perc_baseline_sd^2) * (baseline_n - 1)) / sum(baseline_n - 1))) %>% 
  mutate(perc_se_baseline = pooled_perc_sd_baseline / sqrt(n_overall)) %>%  # now also do this for untransformed variable
  mutate(baseline_overall = sum(baseline_mean * baseline_n) / sum(baseline_n)) %>% 
  mutate(pooled_sd_baseline = sqrt(sum((baseline_sd^2) * (baseline_n - 1)) / sum(baseline_n - 1))) %>% 
  mutate(se_baseline = pooled_sd_baseline / sqrt(n_overall))

@@

# Clinical sample

@@

#B. Sex meta-analysis

@@

df_for_sex_metan <- df_for_sex_metan %>% 
  filter(percent_women != 100) # filter out studies with only women

@@

# C. Age meta-analysis--------------------

@@

For this analysis, we excluded the two psychotherapy trials which included entirely female samples (Moeini, 2019; Shomaker, 2016; see @tbl-all-studies-percfemale-results in Supplement for results with full sample). As can be seen in @tbl-baseline_results, psychotherapy trials featured a significantly higher percentage of females when compared to medication trials. On average, samples were `r round(results_met_perc_women_overall$TE_random[results_met_perc_women_overall$subgroup == 1][[2]],2)`% (*SE* = `r round(results_met_perc_women_overall$seTE_random[results_met_perc_women_overall$subgroup == 1][[2]],2)`) female across psychotherapy trials and `r round(results_met_perc_women_overall$TE_random[results_met_perc_women_overall$subgroup == 0][[2]],2)`% (*SE* = `r round(results_met_perc_women_overall$seTE_random[results_met_perc_women_overall$subgroup == 0][[2]],2)`) female across medication trials. Excluding sub-clinical and waitlist control studies yielded similar results.

@@

We conducted a series of sensitivity analyses (all figures in the Supplement). Excluding waitlist control studies (see @fig-plot-means-no-wl) and sub-clinical studies (see @fig-plot-means-clin) yielded a pattern of results very similar to the overall analyses. Next, we examined the data including only those studies that used the CDRS (see @fig-plot-means-cdrs) or the HAMD (see @fig-plot-means-hamd). Medication control and psychotherapy control conditions remained significantly different, though the small number of studies resulted in less precise estimates of the SMDs. Finally, we showed that different values for the pre-post measure correlation had minimal effect on the estimated outcomes (see @fig-stab-sims).

@@

#Repeating meta-analyses on sample characteristics including studies with females only

#An issue that arise when trying to include studies with 100% female is that their SE ends up being 0, which is treated as NA in the metagen function. One way to circumvent this is to apply a continuity correction by adding a constant (e.g 0.5) to percent female. These studies will now be treated as having 99.5% females.

#overall sample------------

df_for_sex_metan_no_excl <- df_baseline_demographics %>% 
  filter(!is.na(percent_women)) %>% #use full dataset to include all studies
  filter(new_study_id != "March, 2004")# removing TADS because it does not technically belong to either the psy or med groups

#using the formula for se (p*(1-p)/n) we add a constant for studies w 100% female

correction_constant <- 0.5

df_for_sex_metan_no_excl <- df_for_sex_metan_no_excl %>% 
  mutate(constant_percent_women = ifelse(percent_women==100,percent_women-correction_constant,percent_women)) %>%
  mutate(product_perc_women=(constant_percent_women/100)*(1-(constant_percent_women/100))) %>%
  mutate(percent_women_std_error = sqrt(product_perc_women / n_overall))

met_perc_women_no_excl  <- metagen(TE = percent_women,
                           seTE = percent_women_std_error,
                           studlab = new_study_id,
                           data = df_for_sex_metan_no_excl,
                           sm = "",
                           fixed = FALSE,
                           random = TRUE,
                           method.tau = "REML",
                           hakn = TRUE,
                           title = "percentage women across studies")

    #Issue: standard error is 0 for 100% female studies so the model     treats them as NA! fix if we want to include all studies in these analyses

met_perc_women_no_excl <- update(met_perc_women_no_excl,
                              subgroup = df_for_sex_metan_no_excl$psy_or_med, subgroup.name = "treatment modality",
                              tau.common = FALSE)

results_met_perc_women_no_excl <- extract_stats(met_perc_women_no_excl)

# Clinical only-------------

df_for_sex_metan_no_excl_clin <- df_for_sex_metan_no_excl %>%
  filter(new_study_id %in% clin_study_ids)

met_perc_women_no_excl_clin <- metagen(TE = percent_women,
                           seTE = percent_women_std_error,
                           studlab = new_study_id,
                           data = df_for_sex_metan_no_excl_clin,
                           sm = "",
                           fixed = FALSE,
                           random = TRUE,
                           method.tau = "REML",
                           hakn = TRUE,
                           title = "percentage women across clinical studies")

met_perc_women_no_excl_clin <- update(met_perc_women_no_excl_clin,
                           subgroup = df_for_sex_metan_no_excl_clin$psy_or_med, subgroup.name = "treatment modality",
                           tau.common = FALSE)

# Apply the function to the meta-analysis result
results_met_perc_women_no_excl_clin <- extract_stats(met_perc_women_no_excl_clin)

# Excluding wl--------------

df_for_sex_metan_no_excl_no_wl <- df_for_sex_metan_no_excl %>%
  filter(new_study_id %in% no_wl_ids)

met_perc_women_no_excl_no_wl <- metagen(TE = percent_women,
                           seTE = percent_women_std_error,
                           studlab = new_study_id,
                           data = df_for_sex_metan_no_excl_no_wl,
                           sm = "",
                           fixed = FALSE,
                           random = TRUE,
                           method.tau = "REML",
                           hakn = TRUE,
                           title = "percentage women across studies with controls that are not wl",
                           subgroup = psy_or_med,
                           tau.common = FALSE)

# Apply the function to the meta-analysis result
results_met_perc_women_no_excl_no_wl <- extract_stats(met_perc_women_no_excl_no_wl)

## Combine results for table

df_baseline_results_no_excl <- rbind(results_met_perc_women_no_excl, results_met_perc_women_no_excl_clin, results_met_perc_women_no_excl_no_wl)

# For row names
df_baseline_results_no_excl <- df_baseline_results_no_excl %>%
  mutate(subgroup = case_when(
    row_number() %in% 1 ~ "Overall",
    row_number() %in% 4 ~ "Excluding subclinical",
    row_number() %in% 7 ~ "Excluding waitlist",
    TRUE ~ if_else(subgroup == 0, "Medication", "Psychotherapy")
  ))

# Assign new column names to the data frame
colnames(df_baseline_results_no_excl) <- new_column_names

# Rounding
df_baseline_results_no_excl <- df_baseline_results_no_excl %>%
  mutate(across(.cols = 2:(ncol(.) - 1), ~ round(., 2))) %>%
  mutate_at(vars(ncol(.)), ~ round(., 3)) %>%
  mutate_all(~if_else(is.na(.), "", as.character(.)))

```{r, warning=FALSE, message = FALSE, echo=FALSE, ft.align="left"}
#| label: tbl-all-studies-percfemale-results
#| tbl-cap: "Percentage female at baseline across medication and psychotherapy studies: Results for sample including all female studies"
#| tbl-cap-location: top

df_baseline_results_no_excl %>% 
  as_flextable() %>% 
  bold(i = c(1, 4, 7), j = NULL, bold = TRUE, part = "body") %>%  # Bold specific rows
  # set_table_properties(layout = "fixed") %>%  # Autofit column widths
  bg(i = 1, bg  = "#F2F2F2", part = "body") %>% # identify category sections
  padding(i = c(2, 3, 5, 6, 8, 9), j = 1, padding.left = 10) %>% # indent some rows
  align(j = 2:ncol(df_baseline_results_no_excl), align = "center", part = "all") %>% 
  bold(part = "header") %>%   # Bold the relevant header rows
  bg(bg = "#F2F2F2", part = "header") %>%   # Add light grey background to header row  
  # italic(i = 2, italic = TRUE, part = "header") %>% 
  border_remove() %>% # this is to remove borders that look strange
  hline(part = "body", border = small_border) %>%  # Add light grey horizontal lines all  rows
  vline(j = 1, part = "body", border = small_border) %>% 
  border_outer(border = small_border, part = "all" ) %>%  # Add outer borders
  hline(part = "header", border = big_border) %>% 
  # padding(i = 2, padding.bottom = 15, part = "header") %>% 
  line_spacing(space = .7, part = "all") %>%
  font(fontname = "Calibri", part = "all") %>%   # Set font to Calibri
  autofit()
